---
title: 'CAMDA 2018: Cancer Data Integration Challenge'
subtitle: '--- **Neuroblastoma: sample analysis** ---'
author: "`r if (exists('yml')) yml$analyst else 'Jim Zhang'`"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    number_sections: no
    toc: yes
---

```{r global_setup, eval=TRUE, include=FALSE}
## Default knitr parameters
knitr::opts_chunk$set(dpi=300, dev=c('png', 'pdf'), echo=FALSE, warning=FALSE, message=FALSE);

require(rmarkdown);
require(knitr); 
require(awsomics);
require(VennDiagram);
```

# Clinical information

<div align='center'>
```{r clinical_info, include=TRUE, fig.width=10, fig.height=4, out.width='1000px'}
cl <- readRDS('../../R/clinical.rds'); 

cnm <- c("Country", "INSS Stage", "HighRisk", "MYCN status", "Sex_Imputed", 
         "Training/Validation set", "Class label", "EFS_bin", "OS_bin", 
         "clinico genetic subgroup");
cnm <- sort(cnm);

par(mfrow=c(2, 5)); 
for (i in 1:length(cnm)) pie(table(cl[, cnm[i]]), col=rainbow(length(unique(cl[, cnm[i]]))),
                             main=cnm[i]); 
```
</div>

<div style="color:darkblue; padding:0 1cm">
**Figure 1A.** Clinical characteristics of the 498 neuroblastoma patients.
</div>

<div align='center'>
```{r clinical_age, include=TRUE, fig.width=6.4, fig.height=4, out.width='640px'}
par(mar=c(4,5,2,2));
hist(as.numeric(cl[, 'Age']), br=100, xlab='', ylab='Number of patients', main='Age');
```
</div>

<div style="color:darkblue; padding:0 3cm">
**Figure 1B.** Clinical characteristics of the 498 neuroblastoma patients: Age distribution.
</div>

# Expression

## PCA 

### GSE49711/RNA-seq {.tabset}

```{r PCA_GSE49711, eval=FALSE, include=FALSE}
dir.create('PCA', showWarnings = FALSE);
e1 <- readRDS('../../R/expression/GSE49711/expr_gene.rds')[, rownames(cl)];
pr <- prcomp(t(e1)); 

fns <- sapply(cnm, function(nm) { 
  print(nm);
  fn <- paste('PCA/GSE49711_', gsub('[ /]', '', nm), '.png', sep='');
  png(fn, w=1200, h=900); 
  cex <- 2.5;
  c <- cl[, nm]; 
  c[is.na(c) | c==''] <- 'Not available';
  col <- rainbow(length(unique(c))); 
  col[col=='#FFFF00FF'] <- '#EEEE00FF'; 
  col <- sub('FF$', 'DD', col); 
  names(col) <- unique(c); 
  col <- col[as.character(c)]; 
  PlotPCA(pr, groups=c, new.window = FALSE, cex=cex, col=col, legend=TRUE); 
  if (dev.cur()!=1) dev.off(); 
  fn; 
}); 
```

```{r PCA_GSE49711_out, include=FALSE}
f <- dir("PCA");
fns <- paste('PCA', f[grep('GSE49711', f)], sep='/');
lns <- paste(paste('####', cnm), '\n\n', paste('![](', fns, ')', sep=''), '\n\n', collapse='');
```

`r lns`

<div style="color:darkblue; padding:0 0.5cm">
**Figure 2A.** PCA plots of the same 498 NB tumors, using data from [GSE49711](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE49711).
</div>

### GSE62564/RNA-seq {.tabset}

```{r PCA_GSE62564, eval=FALSE, include=FALSE}
e2 <- readRDS('../../R/expression/GSE62564/expr.rds')[, rownames(cl)];
pr <- prcomp(t(e2)); 

fns <- sapply(cnm, function(nm) { 
  print(nm);
  fn <- paste('PCA/GSE62564_', gsub('[ /]', '', nm), '.png', sep='');
  png(fn, w=1200, h=900); 
  cex <- 2.5;
  c <- cl[, nm]; 
  c[is.na(c) | c==''] <- 'Not available';
  col <- rainbow(length(unique(c))); 
  col[col=='#FFFF00FF'] <- '#EEEE00FF'; 
  col <- sub('FF$', 'DD', col); 
  names(col) <- unique(c); 
  col <- col[as.character(c)]; 
  PlotPCA(pr, groups=c, new.window = FALSE, cex=cex, col=col, legend=TRUE); 
  if (dev.cur()!=1) dev.off(); 
  fn; 
}); 
lns <- paste(paste('####', cnm), '\n\n', 
             paste('![](', fns, ')', sep=''), '\n\n', collapse='');
```

```{r PCA_GSE62564_out, include=FALSE}
f <- dir("PCA");
fns <- paste('PCA', f[grep('GSE62564', f)], sep='/');
lns <- paste(paste('####', cnm), '\n\n', paste('![](', fns, ')', sep=''), '\n\n', collapse='');
```

`r lns`

<div style="color:darkblue; padding:0 0.5cm">
**Figure 2B.** PCA plots of the same 498 NB tumors, using data form [GSE62564](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE62564).
</div>

### GSE49710/Microarray {.tabset}

```{r PCA_GSE49710, eval=FALSE, include=FALSE}
e3 <- readRDS('../../R/expression/GSE49710/expr.rds')[, rownames(cl)];
pr <- prcomp(t(e3)); 

fns <- sapply(cnm, function(nm) { 
  print(nm);
  fn <- paste('PCA/GSE49710_', gsub('[ /]', '', nm), '.png', sep='');
  png(fn, w=1200, h=900); 
  cex <- 2.5;
  c <- cl[, nm]; 
  c[is.na(c) | c==''] <- 'Not available';
  col <- rainbow(length(unique(c))); 
  col[col=='#FFFF00FF'] <- '#EEEE00FF'; 
  col <- sub('FF$', 'DD', col); 
  names(col) <- unique(c); 
  col <- col[as.character(c)]; 
  PlotPCA(pr, groups=c, new.window = FALSE, cex=cex, col=col, legend=TRUE); 
  if (dev.cur()!=1) dev.off(); 
  fn; 
}); 
```

```{r PCA_GSE49710_out, include=FALSE}
f <- dir("PCA");
fns <- paste('PCA', f[grep('GSE49710', f)], sep='/');
lns <- paste(paste('####', cnm), '\n\n', paste('![](', fns, ')', sep=''), '\n\n', collapse='');
```

`r lns`

<div style="color:darkblue; padding:0 0.5cm">
**Figure 2C.** PCA plots of the same 498 NB tumors, using data from [GSE49710](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE49710).
</div>

# CNV

## GEO series and platforms

```{r cnv, include=FALSE}
cnv <- readRDS('../../R/aCGH/cnv_all.rds');
gse <- sapply(strsplit(names(cnv), '_'), function(x) x[1]); 
plt <- sapply(strsplit(names(cnv), '_'), function(x) x[2]);
plt[plt=='GPL16865'] <- 'GPL2879';

x <- rep(gse, lapply(cnv, ncol));
y <- rep(plt, lapply(cnv, ncol));
z <- xtabs(~x+y);
z <- as.matrix(z);

m <- readRDS('../../R/aCGH/cnv_meta.rds');
```

<div style="color:darkblue; padding:0 4cm">
**Table 1** Summary of CNV platforms and series.
</div>

<div align='center', style="padding:0 4cm">
`r kable(m, row.names=FALSE)`
</div>

<div style="color:darkblue; padding:0 5cm">
**Table 2** Numbers of NB tumors measured by CNV platforms and series. 
</div>

<div align='center', style="padding:0 6cm">
`r kable(z)`
</div>

<div align='center'>
```{r cnv_series, include=TRUE, fig.height=3.2, fig.width=4.8, out.width='480px'}
tid <- lapply(unique(gse), function(g) lapply(cnv[gse==g], colnames));
tid <- lapply(tid, unlist);
tid <- lapply(tid, as.vector);
tid <- lapply(tid, unique); 
num0 <- sapply(tid, length); 

ind1 <- t(combn(1:4, 2));
num1 <- apply(ind1, 1, function(i) length(Reduce('intersect', tid[i]))); 

ind2 <- lapply(4:1, function(i) (1:4)[-i]); 
num2 <- sapply(ind2, function(i) length(Reduce('intersect', tid[i]))); 

num3 <- length(Reduce('intersect', tid));

par1 <- as.list(c(num0, num1, num2, num3));
par1$category <- unique(gse);
par1$fill <- rainbow(length(unique(gse)));

par(mfrow=c(1, 2)); 
do.call('draw.quad.venn', par1) -> x; 
```
</div>

<div style="color:darkblue; padding:0 5cm">
**Figure 3A.** Overlapping of CNV samples measured by GEO series.
</div>

<div align='center'>
```{r cnv_platform, include=TRUE, fig.height=3.2, fig.width=4.8, out.width='480px'}
tid <- lapply(unique(plt), function(g) lapply(cnv[plt==g], colnames));
tid <- lapply(tid, unlist);
tid <- lapply(tid, as.vector);
tid <- lapply(tid, unique); 
num0 <- sapply(tid, length); 

ind1 <- t(combn(1:4, 2));
num1 <- apply(ind1, 1, function(i) length(Reduce('intersect', tid[i]))); 

ind2 <- lapply(4:1, function(i) (1:4)[-i]); 
num2 <- sapply(ind2, function(i) length(Reduce('intersect', tid[i]))); 

num3 <- length(Reduce('intersect', tid));

par2 <- as.list(c(num0, num1, num2, num3));
par2$category <- unique(plt);
par2$fill <- rainbow(length(unique(plt)));

do.call('draw.quad.venn', par2) -> x; 
```
</div>

<div style="color:darkblue; padding:0 5cm">
**Figure 3B.** Overlapping of CNV samples measured by GEO platforms.
</div>


## PCA

### GSE45480_GPL4093 {.tabset}

```{r PCA_GSE45480_GPL4093, eval=FALSE, include=FALSE}
c1 <- cnv[['GSE45480_GPL4093']];
pr <- prcomp(t(c1)); 

fns <- sapply(cnm, function(nm) { 
  print(nm);
  fn <- paste('PCA/GSE45480_GPL4093_', gsub('[ /]', '', nm), '.png', sep='');
  png(fn, w=1200, h=900); 
  cex <- 2.5;
  c <- cl[colnames(c1), nm]; 
  c[is.na(c) | c==''] <- 'Not available';
  col <- rainbow(length(unique(c))); 
  col[col=='#FFFF00FF'] <- '#EEEE00FF'; 
  col <- sub('FF$', 'DD', col); 
  names(col) <- unique(c); 
  col <- col[as.character(c)]; 
  PlotPCA(pr, groups=c, new.window = FALSE, cex=cex, col=col, legend=TRUE); 
  if (dev.cur()!=1) dev.off(); 
  fn; 
}); 
```

```{r PCA_GSE45480_GPL4093_out, include=FALSE}
f <- dir("PCA");
fns <- paste('PCA', f[grep('GSE45480_GPL4093', f)], sep='/');
lns <- paste(paste('####', cnm), '\n\n', paste('![](', fns, ')', sep=''), '\n\n', collapse='');
```

`r lns`

<div style="color:darkblue; padding:0 0.5cm">
**Figure 4A.** PCA plots of `r ncol(cnv[['GSE45480_GPL4093']])` NB tumors, using data from GEO series GSE45480 on platform GPL4093.
</div>

### GSE45480_GPL2873 {.tabset}

```{r PCA_GSE45480_GPL2873, eval=FALSE, include=FALSE}
c2 <- cnv[['GSE45480_GPL2873']];
pr <- prcomp(t(c2)); 

fns <- sapply(cnm, function(nm) { 
  print(nm);
  fn <- paste('PCA/GSE45480_GPL2873_', gsub('[ /]', '', nm), '.png', sep='');
  png(fn, w=1200, h=900); 
  cex <- 2.5;
  c <- cl[colnames(c2), nm]; 
  c[is.na(c) | c==''] <- 'Not available';
  col <- rainbow(length(unique(c))); 
  col[col=='#FFFF00FF'] <- '#EEEE00FF'; 
  col <- sub('FF$', 'DD', col); 
  names(col) <- unique(c); 
  col <- col[as.character(c)]; 
  PlotPCA(pr, groups=c, new.window = FALSE, cex=cex, col=col, legend=TRUE); 
  if (dev.cur()!=1) dev.off(); 
  fn; 
}); 
```

```{r PCA_GSE45480_GPL2873_out, include=FALSE}
f <- dir("PCA");
fns <- paste('PCA', f[grep('GSE45480_GPL2873', f)], sep='/');
lns <- paste(paste('####', cnm), '\n\n', paste('![](', fns, ')', sep=''), '\n\n', collapse='');
```

`r lns`

<div style="color:darkblue; padding:0 0.5cm">
**Figure 4B.** PCA plots of `r ncol(cnv[['GSE45480_GPL2873']])` NB tumors, using data from GEO series GSE45480 on platform GPL4093.
</div>

### GSE25771_GPL2873 {.tabset}

```{r PCA_GSE25771_GPL2873, eval=FALSE, include=FALSE}
c3 <- cnv[['GSE25771_GPL2873']];
pr <- prcomp(t(c3[!is.na(rowMeans(c3)), ])); 

fns <- sapply(cnm, function(nm) { 
  print(nm);
  fn <- paste('PCA/GSE25771_GPL2873_', gsub('[ /]', '', nm), '.png', sep='');
  png(fn, w=1200, h=900); 
  cex <- 2.5;
  c <- cl[colnames(c3), nm]; 
  c[is.na(c) | c==''] <- 'Not available';
  col <- rainbow(length(unique(c))); 
  col[col=='#FFFF00FF'] <- '#EEEE00FF'; 
  col <- sub('FF$', 'DD', col); 
  names(col) <- unique(c); 
  col <- col[as.character(c)]; 
  PlotPCA(pr, groups=c, new.window = FALSE, cex=cex, col=col, legend=TRUE); 
  if (dev.cur()!=1) dev.off(); 
  fn; 
}); 
```

```{r PCA_GSE25771_GPL2873_out, include=FALSE}
f <- dir("PCA");
fns <- paste('PCA', f[grep('GSE25771_GPL2873', f)], sep='/');
lns <- paste(paste('####', cnm), '\n\n', paste('![](', fns, ')', sep=''), '\n\n', collapse='');
```

`r lns`

<div style="color:darkblue; padding:0 0.5cm">
**Figure 4B.** PCA plots of `r ncol(cnv[['GSE25771_GPL2873']])` NB tumors, using data from GEO series GSE25771 on platform GPL2873.
</div>



## Data distribution

<div align='center'>
```{r, cnv_dist, include=TRUE, fig.width=20, fig.height=12, out.width='1000px'}
cnt <- lapply(cnv, function(c) table(round(as.vector(c), 2)));

par(mfrow=c(2, 5), mar=c(5,5,3,2)); 
for (i in 1:length(cnt)) {
  x <- cnt[[i]]; 
  plot(as.numeric(names(x)), as.integer(x), type='h', log='y', col='#88888888', 
       xlab="CNV measurements", ylab="Number of calls", main=names(cnt)[i]);
}; 
```
</div>
<div style="color:darkblue; padding:0 0.5cm">
**Figure 5.** Distribution of CNV measurements, by GEO series and platforms. 
</div>

# Session information

```{r session_info, echo=FALSE}
sessionInfo(); 
```

***
_END OF DOCUMENT_

